Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motions Characteristic to Vertical Axis Turbines

### 2016-03-06 paper.tex
new script with new figures
Clean up abstract wording a bit

    Vertical Axis Wind Turbines (VAWTs) have been used for extracting of energy
    either from sea or wind. A simpler design, low cost of maintenance and the
    capability of working independent from the direction of the flow are some of
    the most important advantages over conventional Horizontal Axis Wind
    Turbines (HAWTs). Complex and unsteady fluid dynamics are not easy to model
    for simulation tools. One of the most relevant phenomena is the dynamic
    stall which is inherently to unsteady conditions, this work is focused in
    its study. Dynamic stall is usually used as a passive control for VAWT
    operating conditions, hence the importance of a good understanding. In this
    study, \textbf{a coupled model is implemented with the open-source CFD
    toolbox OpenFOAM for solving the governing equations; an actuator line model
    is used for solving velocity flow field and a dynamic stall model for force
    interactions.} Force coefficients obtained from the model are validated with
    experimental data of pithing airfoil in similar operating conditions as
    H-rotor type. Numerical results show reasonable agreement with experimental
    data for pitching motion.

@@

\begin{figure}[h]
\begin{minipage}{18pc}
\includegraphics[width=18pc]{Fig1.png}
\end{minipage}\hspace{2pc}%
\begin{minipage}{18pc}
\includegraphics[width=18pc]{Fig2.png}
\end{minipage}
\caption{\label{fig1}Normal (left) and tangential (right) force coefficients during pitching motions of NACA0021 airfoil with a maximum amplitude of 17.4\degree\ (analog to $\lambda = 3.34$).}
\end{figure}

\begin{figure}[h]
\begin{minipage}{18pc}
\includegraphics[width=18pc]{Fig3.png}
\end{minipage}\hspace{2pc}%
\begin{minipage}{18pc}
\includegraphics[width=18pc]{Fig4.png}
\end{minipage}
\caption{\label{fig2}Normal (left) and tangential (right) force coefficients during pitching motions of NACA0021 airfoil with a maximum amplitude of 22.6\degree\ (analog to $\lambda = 2.60$).}
\end{figure}

>>>>

    Vertical axis turbines (VATs) can be used to extract renewable energy from
    wind or water flows. A simpler design, low cost of maintenance, and the
    ability to accept flow from all directions perpendiculat to the rotor axis
    are some of the most important advantages over conventional horizontal axis
    turbines (HATs). However, VATs encounter complex and unsteady fluid
    dynamics, which present significant modeling challenges. One of the most
    relevant phenomena is dynamic stall, which is caused by the unsteady
    variation of angle of attack throughout the blade rotation, and is the focus
    of the present study. Dynamic stall is usually used as a passive control for
    VAT operating conditions, hence the importance of predicting its effects. In
    this study, \textbf{a coupled model is implemented with the open-source CFD
    toolbox OpenFOAM for solving the Navier--Stokes equations, where an actuator
    line model and dynamic stall model are used compute the blade loading and
    body force.} Force coefficients obtained from the model are validated with
    experimental data of pithing airfoil in similar operating conditions as
    H-rotor type. Numerical results show reasonable agreement with experimental
    data for pitching motion.

@@

\begin{figure}[h]
\begin{minipage}{18pc}
%\includegraphics[width=18pc]{Fig1.png}
\resizebox{\columnwidth}{!}{\input{CN1}}
\end{minipage}\hspace{2pc}%
\begin{minipage}{18pc}
%\includegraphics[width=18pc]{Fig2.png}
\resizebox{\columnwidth}{!}{\input{CT1}}
\end{minipage}
\caption{\label{fig1}Normal (left) and tangential (right) force coefficients during pitching motions of NACA0021 airfoil with a maximum amplitude of 17.4\degree\ (analog to $\lambda = 3.34$).}
\end{figure}

\begin{figure}[h]
\begin{minipage}{18pc}
%\includegraphics[width=18pc]{Fig3.png}
\resizebox{\columnwidth}{!}{\input{CN2}}
\end{minipage}\hspace{2pc}%
\begin{minipage}{18pc}
%\includegraphics[width=18pc]{Fig4.png}
\resizebox{\columnwidth}{!}{\input{CT2}}
\end{minipage}
\caption{\label{fig2}Normal (left) and tangential (right) force coefficients during pitching motions of NACA0021 airfoil with a maximum amplitude of 22.6\degree\ (analog to $\lambda = 2.60$).}
\end{figure}
